---
title: "Practicum 1"
author: "Chenyao Xiao, Yicheng Zhang and Isabella Motha"
date: "2022-10-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(skimr)
library(psych)
library(janitor)
library(XML)
library(xml2)
library(RCurl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
library(readr)
library(tibble)
```

## Part 1

### 1
```{r}
# Create 4 variables
doctor_type <- c("PCP", "Psychiatrist", "Surgeon", "Anesthesia")
doctor_lastname <- c("Smith", "Dame", "Jones", "Zayas")
location <- c("MA", "ME", "NH", "VT")
AVG_Rating <- c(7,9,8,9)
# Join the variables to create a data frame
doctor_df <- data.frame(doctor_type, doctor_lastname, location, AVG_Rating)
doctor_df
```

### 2
```{r}
# Select row 1 in column 2
doctor_df[1,2]
# Select rows 2 through 4
doctor_df[2:4,]
# Select the last column
doctor_df[, ncol(doctor_df)]
```

### 3
```{r}
mtcars$vs <- as.factor(mtcars$vs)
ggplot(data = mtcars) +
geom_point(mapping = aes(x = disp, y = mpg, color = vs))+
  labs(title= "The Relationship between Displacement and Mile per Gallon",
       x="Displacement (cu.in)",  y="Miles per gallon", caption="Measure relationship 
       between fuel efficiency and engine size in mtcars")+
theme_classic()
```
Here we choose two variables: Miles per gallon to measure fuel efficiency and Displacement
to measure engine size. Also, we choose a color scheme of vs(engine shape). Red spots
represent V-shaped engine, while green stands for straight engines.

The scatter plot shows a negative relationship between engine displacement(cu.in) 
and fuel efficiency(miles per gallon). 
Basically, this is because bigger engine cars use more fuel and can reduce fuel efficiency. 
The color in this plot reveals that V-shaped engines with less efficiency use 
more fuel since they have bigger engine size.

### 4
```{r}
# Perform a summary statistic on the dataset
summary(mtcars)
# calculate the pearson coefficient of the correlation R
cor(mtcars$mpg, mtcars$cyl,method = "pearson", use = "complete.obs")
```
Since we can get a negative correlation between displacement and mpg, we can pick
other values like cylinders. More cylinders tend to need large engine size. Thus we
assume that the variable cyl and mpg are negatively correlated.
The correlation value -0.852 with pearson method proved that my assumption is correct. 
The more cylinders a car has, the lower mpg it might achieve. The purpose of this coefficient 
is to offer us if these two variable has association (the correct coefficient should be -1 to 1) 
and how is their association (positive or negative).

## Part 2

### 1
```{r,message=FALSE}
# Load the XML data directly from the URL 
# it might take a while to run this chunk of code!:
fileurl <- "https://data.ny.gov/api/views/ngbt-9rwf/rows.xml"
download.file(fileurl, destfile = "rows.xml")
# parse into a data frame 
rows_xml <- xmlParse("rows.xml")
ny.cd_df <- xmlToDataFrame(nodes=getNodeSet(rows_xml, "//response/row/row"))
```

### 2
After loading the data, we can check the structure of the data frame.
```{r}
dim(ny.cd_df)
glimpse(ny.cd_df) 
summary(ny.cd_df)
```
This data frame contains 7 variables and 99367 observations. According to the data dictionary,
some variables needs to be cleaned as other formats. The data frame doesn't have 
NAs.
```{r}
# Change year and admissions to numeric
ny.cd_df2 <-ny.cd_df %>%
  mutate(across(c(year,admissions),as.numeric))
# Find NAs
sapply(ny.cd_df2,function(x) sum(is.na(x)))
```

Outliers increase the variability in the data, which decreases statistical power. 
Consequently, excluding outliers can cause our results to become statistically 
significant. For admission numbers here, we need to remove outliers. When we expect 
our data has a normal distribution, Z-scores may help us to identify how odd the 
real observation is. 

The degree of standard deviations around the mean decrease for each value is represented 
by the Z-score. The Z score equals to zero refers to the mean value, and in order 
to calculate it, we need to take the original admissions, calculate the mean and 
the use  it to minus origin admission value, then divide by the SD to get the 
Z-score of the admission. 

A huge the distance between an admission's Z-score and zero can indicate they are 
actually extravagant. Z ± 3 is a widely accepted cut-off point for identifying 
outliers, and we will apply the same logistics in our computation.

```{r} 
# Remove outliers using z score
# When z > 3, it means the data is more than
# 3 times standard deviation from the mean, which is considered an outlier
Adm <- ny.cd_df2$admissions
mean_Adm <- mean(Adm)
Std.Dev <- sd(Adm)
ny.cd_df2$z <-(mean_Adm - Adm)/Std.Dev
ny.cd_df2$z <-abs(ny.cd_df2$z)
ny.cd_df2 <- ny.cd_df2 %>%
  filter(z<=3) 
```
Then we can remove column z.
```{r}
ny.cd_df2 <- ny.cd_df2[,-8]
```
We can see that 97450 rows left after removing outliers.

### 3
Next we prepare four tibbles.
```{r}
# tibble 1: county

county <- tibble(county_code = c("AL","CA","CN","DE","FR","LE","MG","ON","OL","NY",
                                 "SL","SY","SV","WR","AG","CY","CL","DU","FU","HE",
                                 "LI","NA","OD","OS","RE","SA","SE","TI","WS","NY",
                                 "CH","CO","ER","GE","JE","MA","NY","OT","OG","NY",
                                 "SC","ST","TO","WA","BM","CM","CR","ES","GR","NY",
                                 "MO","NI","OR","PU","RO","SH","SU","UL","WE","WY",
                                 "YA"),
                 county_of_program_location = c("Albany","Cattaraugus","Chenango","Delaware","Franklin",
                                                "Lewis","Montgomery","Oneida","Orleans","Queens",
                                                "Saint Lawrence","Schuyler","Sullivan","Warren",
                                                "Allegany","Cayuga","Clinton","Dutchess",
                                          "Fulton","Herkimer","Livingston","Nassau","Onondaga",
                                          "Oswego","Rensselaer","Saratoga","Seneca","Tioga",
                                          "Washington","Bronx","Chautauqua","Columbia","Erie",
                                          "Genesee","Jefferson","Madison","New York","Ontario",
                                          "Otsego","Richmond","Schenectady","Steuben","Tompkins",
                                          "Wayne","Broome","Chemung","Cortland","Essex","Greene",
                                          "Kings","Monroe","Niagara","Orange","Putnam",
                                          "Rockland","Schoharie","Suffolk","Ulster","Westchester",
                                          "Wyoming","Yates"))

# make sure they are all unique:

county %>% distinct(county_code,.keep_all = TRUE)

```

```{r}
# tibble 2 : program_category:

program_category <-tibble(program_code = c("CR", "IN","OTP", "RE", "OU", "SP"),
                          program_category = c("Crisis", "Inpatient", "Opioid Treatment Program",
                                               "Residential", "Outpatient", "Specialized"))
program_category %>% distinct(program_code,.keep_all = TRUE)
```
```{r}
# tibble 3:primary_substance_group

primary_substance_group <- tibble(substance_code=c("H","A","AO","C","M","N","O"),
                                  primary_substance_group= c("Heroin", "Alcohol",
                                  "All Others", "Cocaine", "Marijuana","None","Other Opioids"))
primary_substance_group %>% distinct(substance_code,.keep_all = TRUE)
```
To get *admissions-data* tibble, we need to merge our tibbles for code into original
data and select columns properly.

```{r}
# tibble 4 admission_data:
# Instead include a column with their respective foreign keys
ny.cd_df3 <-left_join(ny.cd_df2,county,by="county_of_program_location")
ny.cd_df3 <-left_join(ny.cd_df3,program_category,by="program_category")
ny.cd_df3 <-left_join(ny.cd_df3,primary_substance_group,by="primary_substance_group")

# Select and rename columns we need
admissions_data <- ny.cd_df3 %>%
                 select(year,county_code, program_code, service_type, age_group,
                        substance_code,admissions) %>%
                 rename("county_of_program_location" = "county_code",
                        "program_category" = "program_code",
                       "primary_substance_group" = "substance_code")

# check our result:

str(admissions_data)
head(admissions_data)
```

### 4

```{r}
annualAdmissions <- function(){
    data_4 <- admissions_data %>%
    select(year,admissions)%>%
    group_by(year) %>%
    summarise(sum_adm=sum(admissions))
    ggplot(data_4, aes(x=year, y=sum_adm))+
    geom_line()+
    scale_x_continuous(breaks=seq(2007, 2021, 1))+
    geom_vline(aes(xintercept=2016),color="grey", linetype="dashed", size=0.5)+
    annotate("text",x=2016,y=193000,hjust=.1,label="Highest: 190094 (Year 2016)")+
    labs(title="Reported admissions in NY state from 2007 to 2021",
         x="Year", y="Number of admissions",
         caption = "Admission numbers in NY state reached the highest number in 2016")+
      theme_classic()+
      theme(plot.title = element_text(hjust = 0.5))
}
annualAdmissions()
```

From this result we can see that the general trend of the admission due to substance 
abusing is decreasing after we arrive the peak point on 2016 with 190094 individual 
admission cases. And then start from 2018 the trend is keeping decrease at we arrive 
the lowest point in 2021, which I believe we shold give credit to the improvement 
of the public health system in recent decades otherwise we may not have a such dramatic 
falling of admission numbers since 2016. 

### 5
To ensure correct merging, we need to check the code for the county as unique.
So we recode the county table here.
```{r}
# Prepare data
# since 5 counties share the same code "NY",
# we want to rename their county code to make sure they are unique:
county_unique <- tibble(county_code = c("AL","CA","CN","DE","FR","LE","MG","ON","OL","NYQ",
                                        "SL","SY","SV","WR","AG","CY","CL","DU","FU","HE",
                                        "LI","NA","OD","OS","RE","SA","SE",
                                        "TI","WS","NYB","CH","CO","ER","GE","JE","MA","NY",
                                        "OT","OG","NYR","SC","ST","TO","WA",
                                        "BM","CM","CR","ES","GR","NYK","MO","NI","OR","PU",
                                        "RO","SH","SU","UL","WE","WY","YA"),
                 County_of_Program_Location = c("Albany","Cattaraugus","Chenango",
                                                "Delaware","Franklin","Lewis","Montgomery",
                                                "Oneida","Orleans","Queens","Saint Lawrence","Schuyler","Sullivan","Warren","Allegany","Cayuga","Clinton","Dutchess",
                                              "Fulton","Herkimer","Livingston","Nassau",
                                              "Onondaga","Oswego","Rensselaer","Saratoga",
                                                "Seneca","Tioga","Washington","Bronx",
                                              "Chautauqua","Columbia","Erie","Genesee",
                                                "Jefferson","Madison","New York","Ontario","Otsego","Richmond","Schenectady","Steuben","Tompkins","Wayne","Broome",
                                                "Chemung","Cortland","Essex","Greene",
                                              "Kings","Monroe","Niagara","Orange",
                                                "Putnam","Rockland","Schoharie","Suffolk",
                                              "Ulster","Westchester","Wyoming","Yates"))
county_unique <- county_unique %>% distinct(county_code,.keep_all = TRUE)
# Redo admission data tibbles
admissions_data2 <- merge(x=county_unique,y=ny.cd_df2,by.x="County_of_Program_Location",
                         by.y = "county_of_program_location",all= T)
admissions_data2 <- merge(x=primary_substance_group, y=admissions_data2, by.x="primary_substance_group",
                         by.y = "primary_substance_group", all= T)
admissions_data2 <- merge(x=program_category, y=admissions_data2, by.x="program_category",
                         by.y="program_category",all = T)
# Select and rename columns we need
admissions_data2 <- admissions_data2 %>%
                 select(year,county_code, program_code, service_type, age_group,
                        substance_code,admissions) %>%
                 rename("county_of_program_location" = "county_code",
                        "program_category" = "program_code",
                       "primary_substance_group" = "substance_code")
```
Basically what we do is display question 3 in unique county codes.

```{r}
# Merge tibbles back
colnames(admissions_data2)[2]<-"county_code"
admissions_data_new<-left_join(admissions_data2,county_unique,by="county_code")

# Group and summarise
admissions_data_new <- admissions_data_new %>% 
  group_by(County_of_Program_Location) %>%
  summarise(adm_number= sum(admissions)) %>%
  mutate(percentage=round(adm_number/sum(adm_number),4)*100) %>%
  top_n(5,percentage)
```

```{r}
# Visualize in a bar chart
ggplot(data=admissions_data_new, aes(x=County_of_Program_Location, y=percentage, 
                                   fill=County_of_Program_Location)) +
  geom_bar(stat="identity",color="black")+
  geom_text(aes(label=percentage), vjust=1.6, size=3)+
  labs(title= "Top 5 counties with their percentage of admissions",
       fill= "Counties",x="Counties", y= "Percentage (%)", 
       caption="Find five counties with highest percentage of admissions")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))
```
The top five counties from 2007 to 2021 are New York(city), Queens, Suffolk, Kings 
and Bronx. New York city has the highest admission percentage of 7.64% and it's 
clearly higher than other counties.From this result, we can see that the new york 
city has the highest admission percentage from 2008-2020 and it is way more higher 
than other counties. The Bronx county has the lowest in these 5 counties and it 
is very close to kings. This shows more admission cases in New York(city), 
but the New York city has large population so this is reasonable.

### 6

```{r, message=FALSE}
# extract all admissions to the various “Rehab” facilities.ny.cd_df3
substance2age_admissions_data <- ny.cd_df3 %>%
                  filter(str_detect(service_type, ".*Rehab.*"))%>%
                  group_by(age_group,primary_substance_group) %>%
                  summarise(substance_admission_total =sum(admissions)) 
                  # new column here refers to the total admissions number
```

```{r}
ggplot(substance2age_admissions_data, aes(x=age_group,y=substance_admission_total,fill=primary_substance_group))+
         geom_col(position=position_dodge())+
  scale_fill_brewer(palette="Paired")+
  labs(fill="Substance Group",
       caption="To identify which substance is the most prominent among each age group",
       x="Age Group", y="Numbers",
       title="Primary Substance Abuse among Different Age Group",
       subtitle="from all rehabilitation facilities")+
  # To display groups clearer, we reformat some x labels
  scale_x_discrete("Age Group",
    labels = c(
      "18 through 24" = "18~24",
      "25 through 34" = "25~34",
      "35 through 44" = "35~44",
      "45 through 54" = "45~54"
    )
  )+ theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))
  
```
From the result, we can see that the marijuana take majority of substance using cases 
for under 18. Drug use is more seen in middle age groups. In short, the prominent 
substance for people under 18 is marijuana, and 18 to 24 is Heroin. For age between 
25 to 34, the prominent substance is heroin, though alcohol also has a high abusing cases. 
For people from 35 to 44 and 45 to 54, the prominent substance is alcohol. 
People older than 55 tend to have more issues with alcohol as well. Nearly all age groups
have large cases of alcohol abuses, and it could be consider as prominent as prominent 
in these age groups. Heroin is widely seen as well but most prominent in age 18 to 24 
and 25 to 34.Marijuana is prominent for people under 18.