---
title: "DA5020.A6.YICHENG.ZHANG"
author: "YICHENG ZHANG"
date: "2022-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(RSQLite)
library(DBI)
library(dplyr)
library(ggplot2)
```

# Question 1

## 1.
```{R}
#sqlite> .open imdb.db
#sqlite> .tables
#movie_info
#sqlite> CREATE TABLE director_info (
#   ...> Director_ID INT PRIMARY KEY,
#   ...> Director_Name VARCHAR NOT NULL
#   ...> );
#sqlite> .tables
#director_info  movie_info
```

We now have two tables in our imdb.db database: director_info and movie_info.

## 2.
```{r}
#sqlite> .separator ','
#sqlite> .import directors.csv director_info
#sqlite> SELECT * FROM director_info LIMIT 5
#   ...> ;
#Director_ID,Director
#144,Frank Darabont
#139,Francis Ford Coppola
#86,Christopher Nolan
#459,Sidney Lumet
#sqlite> .save imdb.db
#sqlite> .tables
#director_info  movie_info
```

By using above steps, we now have director_info table with directors.csv imported.

# Question 2

## 1.
```{r, warning=FALSE, message=FALSE}
# we connect to the db first:
# if you want to test this code in your local machine
# pls make sure you modify the path below!!:
db<-dbConnect(SQLite(), dbname="D:/sqlite-tools-win32-x86-3390400/imdb.db")
# see if we have the correct files
dbListTables(db)
# create query
sqlcmd1<-"SELECT * FROM movie_info"
sqlcmd2<-"SELECT * FROM director_info"
# pass query to the sql cmd
rs<-dbGetQuery(db,sqlcmd1)
rs2<-dbGetQuery(db,sqlcmd2)
# count rows:
nrow(rs)
# remove one extra row here since the import of csv will includer headers
rs2<- rs2[-1,]
nrow(rs2)
```

## 2.
```{r, warning=FALSE}
# we send query first:
sqlcmdq2<-"SELECT * FROM movie_info WHERE Release_Year BETWEEN 2010 AND 2020"
movie_2010_to_2020<-dbGetQuery(db,sqlcmdq2)

# count movies in this period:
sqlcmd5<-"SELECT COUNT(*) FROM movie_info WHERE Release_Year BETWEEN 2010 AND 2020 GROUP BY Release_Year"
count_movie<-dbGetQuery(db,sqlcmd5)

# create year column
count_movie$Year<-c(2010:2020)

# visualize the result
ggplot(count_movie, aes(x = Year, y = count_movie[,1], label=Year,fill =Year, color =Year)) +
    geom_line() +
    annotate(geom="text", x=2014, y=32, 
             label="Most movies released") +
    scale_x_continuous(breaks=seq(2010,2020,by=1))+
    ylab("numbers of movies realsed")+
    labs(title = "Movies released from 2010-2020",
        caption = "data from canvas csv file")+
  theme(plot.title = element_text(hjust = 0.5))

```

From above result, we can see that most movies released in 2014. 

## 3.
```{r, warning=FALSE}
sqlcmd3<-"SELECT MAX(IMDB_Rating), MIN(IMDB_Rating),AVG(IMDB_Rating)  FROM movie_info WHERE Genre LIKE '%Action%'"
q3<-dbSendQuery(db,sqlcmd3)
rs_action<-dbFetch(q3)
# here is the max rating:
rs_action[1,1]
# mean rating
rs_action[1,3]
# min rating
rs_action[1,2]
```

The max rating is 9, mean rating is 7.948677 and min rating is 7.6. 

## 4.
```{r,warning=FALSE}
sqlcmd4<-"SELECT Series_Title, Genre, Gross FROM movie_info WHERE GROSS !='NA' ORDER BY Gross DESC LIMIT 25"
top_25_movies<-dbGetQuery(db,sqlcmd4)
top_25_movies
```

Movies listed in the table are 25 movies with highest grossing value.

## 5.
```{R}
# Send query and get the result:
sqlcmdQ5 <- "SELECT director_info.Director_Name, SUM(Gross) AS 'Total' FROM movie_info LEFT JOIN director_info ON director_info.Director_ID = movie_info.Director_ID GROUP BY movie_info.Director_ID ORDER BY Total DESC LIMIT 25"
Best_directors<- dbGetQuery(db, sqlcmdQ5)

# tableview of our query result:
Best_directors

# visualization:
visual_result2 <- ggplot(Best_directors, aes(x=Director_Name,y=Total,label=Total,fill=Director_Name)) +
  geom_bar(stat='identity') +
    ylab("Grossing total")+
    labs(title = "Top 25 directors with highest Grossing total",
        caption = "data from canvas csv file")+
  theme(plot.title = element_text(hjust = 0.5))

# check visual result: 

visual_result2
```

From the result we can tell that Steven Spielberg has the total highest-grossing movies(His total Gross is 2478133165).

## 6.
```{r, warning=FALSE}
verifyDirector <- function(x){
  sqlcmd6 <- "SELECT * FROM director_info WHERE Director_Name = :x"
  send_request <- dbSendQuery(db, sqlcmd6)
  result<- dbFetch(dbBind(send_request, list(x=x)))
  if(nrow(result)==0){
    sprintf("This director ( %s ) doesn't exist in our database", x)
  } 
  else{
    sprintf("This director ( %s ) exists in our database", x)
  }
}

# check our result:
verifyDirector("Christopher Nolan")
verifyDirector("Naruto")
```
