---
title: "DA5020.Practicum 2"
author: "Chenyao Xiao, Yicheng Zhang and Isabella Motha"
date: "2022-11-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(RSQLite)
library(knitr)
```

## Part 1 Customer Database Analysis

Here we need to create a new SQLite database.
```{r}
mydb1 <- dbConnect(RSQLite::SQLite(), "my-db1.sqlite")
```

### 1

The commands for importing data to the database are listed in the sql file. 
<!--
.open "my-db1.sqlite"
CREATE TABLE customers(
CustomerID INT PRIMARY KEY, 
FirstName TEXT, LastName TEXT,
StreetAddress TEXT, City TEXT,
State TEXT, ZipeCode, Telephone,
Purchases_Total NUMERIC);
.mode csv
.import customers.csv temp
INSERT INTO customers SELECT * FROM temp;
DROP TABLE temp;

CREATE TABLE orders(
OrderID INT PRIMARY KEY,
CustomerID, SKU TEXT,
Description TEXT, Cost NUMERIC,
Year_Purchase NUMERIC);
.import orders.csv temp
INSERT INTO orders SELECT * FROM temp;
DROP TABLE temp;

CREATE TABLE Router_Info(
RMAID INT PRIMARY KEY, 
OrderID, Status TEXT,
Step TEXT, Reason TEXT,
CustomerID);
.import Router_Info.csv Router_Info
DELETE
FROM
    Router_Info
WHERE
    RMAID = 'RMAID';
-->

Here we found the headers are duplicated, thus we delete the first row to get the
right Router_Info table.
```{r}
dbListTables(mydb1)
```

### 2

We can check the data by selecting first five rows from each table. 
```{r}
dbGetQuery(mydb1, 'SELECT * FROM customers LIMIT 5;')
dbGetQuery(mydb1, 'SELECT * FROM orders LIMIT 5;')
dbGetQuery(mydb1, 'SELECT * FROM Router_Info LIMIT 5;')
```

By observing three tables, we can tell that they all have 499 observations and unique 
primary keys. Since they all have customerID ,OrderID and RMAID in its own table 
so each row of data should refer to a unique entity with its related in formations. 
We should be able to join all these three tables to find out the complete information 
about each observation. The primary key for customers table is customerID; for 
order table is OrderID; and for Router_info table is RMAID.The three tables contains 
different information are connected by customerID. The table orders and Router_info 
are connected by customerID and orderID. If we want to join them we could do like this:

```{r} 
# we join 3 tables like this: 
sql<-"SELECT * FROM orders JOIN customers ON orders.CustomerID = customers.CustomerID 
JOIN Router_info ON Router_info.CustomerID = orders.CustomerID;"
joinresult<-dbGetQuery(mydb1,sql)
# check result:
str(joinresult)
```

### 3

```{r}
sql_cd1.3 ="SELECT COUNT(CustomerID) FROM orders WHERE COST > 200;"
dbGetQuery(mydb1, sql_cd1.3)
```
We create a query to get that 177 customers spent more than 200 dollars on the orders.

### 4

```{r}
sql_cd1.4 ="SELECT *
FROM customers
JOIN orders ON orders.CustomerID = customers.CustomerID
JOIN Router_Info ON Router_Info.CustomerID = orders.CustomerID;"
df_join <-dbGetQuery(mydb1,sql_cd1.4)
head(df_join)
str(df_join)
```
We can first join two tables of customers and orders by CustomerID and then join Router_Info
table as same. The data frame still contains 499 observations in total but 21 variables now, 
we use direct join here. 

### 5

```{r}
sql_cd1.5 ="SELECT orders.CustomerID, orders.OrderID, Description, 
Purchases_Total, Year_Purchase
FROM customers
LEFT JOIN orders ON orders.CustomerID = customers.CustomerID
LEFT JOIN Router_Info ON Router_Info.CustomerID = orders.CustomerID;"
df_leftjoin <-dbGetQuery(mydb1,sql_cd1.5)
str(df_leftjoin)
```
In this query, we select some basic columns from the tables using left join.

### 6 

The following SQL statement lists the number of customers in each city and their 
average cost total, sorted low to high
(Only include cities with more than 8 customers):
```{r}
sql_cd1.6 ="SELECT COUNT(Router_Info.CustomerID), City, AVG(Cost)
FROM customers
JOIN orders ON orders.CustomerID = customers.CustomerID
JOIN Router_Info ON Router_Info.CustomerID = orders.CustomerID
GROUP BY City
HAVING COUNT(Router_Info.CustomerID) > 8
ORDER BY COUNT(Router_Info.CustomerID);"
df_bycity <- dbGetQuery(mydb1,sql_cd1.6)
kable(df_bycity)
```
We can see that there are 6 cities meeting our query standard. Arlington had most
customers of 15, but Kansas had the highest avarage cost over 180 dollars. 

### 7 

```{r}
sql_cd1.7 ="SELECT Count (SKU) AS 'total_SKU', 
COUNT ( DISTINCT Description ) AS 'total_description_types' 
FROM orders 
WHERE Cost>100;"
dbGetQuery(mydb1,sql_cd1.7)
```
We used the count and count distinctly to know that, among all of 
orders, those orders with cost more than 100, there are 8 distinct 
descriptions types and 340 total SKU exit in the order table.

### 8 

```{r}
sql_cd1.8 = "SELECT Year_Purchase, MIN(Cost), MAX(Cost), SUM(Cost), 
COUNT(OrderID) AS order_numbers
FROM orders
GROUP BY Year_Purchase
ORDER BY COUNT(OrderID) DESC;"
df_yearorders <- dbGetQuery(mydb1,sql_cd1.8)
kable(df_yearorders)
```
The query gets a new table showing us order cost information each year, including
minimum cost, maximum cost, sum and the order numbers. The table is display 
by descending order numbers. The year 2021 has the most orders.

### 9 

```{r}
sql_cd1.9 = "SELECT FirstName, LastName,Status, Step,Reason, Purchases_Total 
FROM customers 
LEFT JOIN Router_info ON Router_info.CustomerID = customers.CustomerID 
WHERE (customers.CustomerID BETWEEN 1001 AND 1021) 
AND (Purchases_Total > 2) 
AND (Purchases_Total < 10)
ORDER by Purchases_Total;"
df_purchase <- dbGetQuery(mydb1,sql_cd1.9)
dim(df_purchase)
kable(head(df_purchase))
```
For this query, we want to know the order status, order steps, reason and
full name of customer who has a customerID between 1001
and 1021 and has a total purchase more than 2 but less than 10. It seems Dean Bernard and 
Brandi Rush have least purchases in these group of customers 

### 10 

```{r}
sql_cd1.10 = "SELECT Year_Purchase, LastName,
COUNT(orders.OrderID) AS order_numbers, Description
FROM customers
LEFT JOIN orders ON orders.CustomerID = customers.CustomerID
WHERE Description LIKE '%Basic Switch%'
AND LastName LIKE 'B%'
GROUP BY Year_Purchase;"
kable(dbGetQuery(mydb1, sql_cd1.10))
```
Here we use wildcard selecting all customers with Last name starting with "B" with 
"Basic Switch" orders and grouping them by Year_purchase.

### 11

```{r}
sql_cd1.11 = "SELECT Year_Purchase, 
COUNT(orders.OrderID) AS order_numbers, AVG(Cost)
FROM orders
LEFT JOIN customers ON customers.CustomerID = orders.CustomerID
WHERE City = 'Washington' OR City = 'Miami'
GROUP BY Year_Purchase;"
kable(dbGetQuery(mydb1, sql_cd1.11))
```
Here we select two cities: Washington and Miami. The query gets order information and
cost average in these cities each year.

```{r}
dbDisconnect(mydb1)
```

## Part 2  HIV Database Analysis

### 1

<!--
.open "my-db2.sqlite"
CREATE TABLE hiv_info (
Entity VARCHAR, Year NUMERIC, 
number_of_people_living_with_HIV INT,
deaths_less_than5 NUMERIC, deaths_more_than70 NUMERIC,
deaths_5to14 NUMERIC, deaths_15to49 NUMERIC,
deaths_50to69 NUMERIC, deaths_total NUMERIC,
new_cases_of_hiv_infection NUMERIC,
number_of_people_infected_with_hiv NUMERIC,
PRIMARY KEY (Entity, Year)
);
CREATE TABLE country_info (
Country VARCHAR, Year NUMERIC, 
Country_Population NUMERIC,
GDP_per_capital NUMERIC,
unemployment_rate FLOAT,
School_enrollment_rate FLOAT,
PRIMARY KEY (Country, Year)
);
.mode csv
.import HIV_info.csv hiv_info
.import country_info.csv country_info
DELETE
FROM
    hiv_info
WHERE
    Year = 'Year';
DELETE
FROM
    country_info
WHERE
    Country = 'Country';
-->

```{r}
mydb2 <- dbConnect(RSQLite::SQLite(), "my-db2.sqlite")
str(dbGetQuery(mydb2, 'SELECT * FROM hiv_info;'))
str(dbGetQuery(mydb2, 'SELECT * FROM country_info;'))
```

There are 44 observations and 11 variables in the HIV_info table and 44 observations
, 6 variables in the country_info table. From these two tables we can see that 
we need to combine country/Entity and year as the primary keys
in order to join them. 

### 2

```{r}
sql_cd2.2 = "SELECT Country, hiv_info.Year, new_cases_of_hiv_infection, 
School_enrollment_rate
FROM hiv_info INNER JOIN country_info ON Entity = Country 
AND hiv_info.Year = country_info.Year
WHERE Country = 'North America';"
df_onlyNA <- dbGetQuery(mydb2,sql_cd2.2)
# Check structure
str(df_onlyNA)
# Check summary statistics
summary(df_onlyNA)
kable(df_onlyNA)
```

For this query, we joined the two table with
```country_info ON Entity = Country AND hiv_info.Year = country_info.Year``` 
to find out the year, new cases of hiv infection and school enrollment rate
of North America.

### 3

The column "number_of_people_living_with_HIV" had a problem that the values were more
than the country population so we need to discard this column. Therefore we choose "number_of_people_infected_with_hiv" minus "deaths_total" to calculate people 
living with hiv. Also here we need to cast the values as float numbers to get decimal outputs.
To calculate the death rate of hiv, we use the deaths_total divided by all people infected.

```{r}
sql_cd2.3 = "SELECT Country, hiv_info.Year, Country_Population, 
CAST((number_of_people_infected_with_hiv - deaths_total) AS FLOAT)*100
/CAST(Country_Population AS FLOAT) AS 'percentage_of_people_living_with_HIV', 
CAST(deaths_total AS FLOAT)*100 /CAST(number_of_people_infected_with_hiv AS FLOAT)
AS 'percentage_deaths_from_HIV'
FROM hiv_info INNER JOIN country_info ON Entity = Country 
AND hiv_info.Year = country_info.Year
WHERE Country = 'North America';"
df_onlyNApercentage <- dbGetQuery(mydb2,sql_cd2.3)
kable(df_onlyNApercentage)
```
By using the ```Cast()``` clause we got the float result of percentage
of people living with hiv and percentage of deaths. We calculate the
percentage of people living with hiv by: 
```(number_of_people_infected_with_hiv - deaths_total)*100/Country_Population```.

We get the death percentage by:```deaths_total*100/number_of_people_infected_with_hiv```.

Still, we only want data from North America this time. The table we got shows that 
from 2007 to 2017, the percentage of people living with HIV
increased from 0.382% to 0.433% while the percentage of deaths decreased from 0.984%
to 0.484%. This is reasonable because more therapies can help reduce deaths while more
people are getting threatened by HIV. 

### 4 

By observing to the ```number_of_people_living_with_HIV``` we know that this is a 
problematic column and we might want to take some extra steps to tidy this column 
before we use it. If we remove the last 3 digits of each number in 
```number_of_people_living_with_HIV``` then it will
have the same number as ```number_of_people_infected_with_hiv```, thus these two 
columns actually refer to same group of people and if we want to use 
```number_of_people_living_with_HIV``` in our query then we need to divide it by 
1000 here:

```{r}
sql_cd2.4="SELECT Country ,SUM(Country_Population) 
AS Sum_country_population, 
CAST(SUM(number_of_people_living_with_HIV/1000*100) AS FLOAT)/CAST(SUM(Country_Population) AS FLOAT) 
AS 'HIV_patient_percentage' 
FROM hiv_info INNER JOIN country_info ON Entity = Country 
AND hiv_info.Year = country_info.Year
WHERE Country != 'North America' GROUP BY Country;"
df_2.4 <-dbGetQuery(mydb2,sql_cd2.4)
kable(df_2.4)
```
We calculated the total population of a country from 2007-2017 and its total HIV 
patient percentage in this period by using ```number_of_people_living_with_HIV/1000``` 
to divide the total population
to get the HIV patient percentage of Bermuda, Canada and United States. And the 
```group by``` clause here is to help us to get the ```HIV_patient_percentage``` 
of these 3 countries individually. The US has the highest HIV patient percentage.

### 5 

#### Query 1

This query helps to investigate relationship between infection rate and unemployment 
rate in Canada. Here we use new cases divided by healthy population 
(```Country_population minus number_of_people_infected_with_hiv```)
to get infection rate. We select data with School_enrollment_rate >99.5 to 
minimise the effect of school enrollment.
```{r}
sql_cd2.5.1 = "SELECT Country, hiv_info.Year, 
CAST(new_cases_of_hiv_infection AS FLOAT)/
CAST(Country_population-number_of_people_infected_with_hiv AS FLOAT) 
AS infection_rate, unemployment_rate
FROM hiv_info INNER JOIN country_info ON Entity = Country 
AND hiv_info.year = country_info.year
WHERE Country = 'Canada' AND School_enrollment_rate >99.5;"
df_5.1 <- dbGetQuery(mydb2, sql_cd2.5.1)
cor(log(df_5.1$infection_rate),df_5.1$unemployment_rate)
model_5.1 <- lm(log(infection_rate)~unemployment_rate,data=df_5.1)
summary(model_5.1)
kable(df_5.1)
```
We can find they are negatively correlated. If unemployment increases, the infection rate
will drop.If there is more people without work, the infection of HIV will go down. 
The p-value 0.296 shows that the correlation is not significant.

#### Query 2

```{r}
sql_cd2.5.2 ="SELECT Country, hiv_info.Year, 
CAST(deaths_15to49 AS FLOAT)/CAST(deaths_total AS FLOAT) 
AS death_15to49_percentage, CAST(deaths_50to69 AS FLOAT)/CAST(deaths_total AS FLOAT) 
AS death_50to69_percentage
FROM hiv_info INNER JOIN country_info ON Entity = Country AND hiv_info.year = country_info.year
WHERE Country = 'United States' AND hiv_info.Year BETWEEN 2011 AND 2015;"
df_5.2 <- dbGetQuery(mydb2, sql_cd2.5.2)
kable(df_5.2)
```
This query suggests that the percentage of deaths between 50 to 69 age is increasing 
from 2011 to 2015, which means the survival time of people infected with HIV could be longer 
as medical conditions gets better.

#### Query 3

Query number 3 we want to calculate the decreasing case of HIV by each year in Bermuda.

```{r}
sql_cd2.5.3 <-"SELECT hiv_info.Year,number_of_people_infected_with_hiv 
AS this_year_HIV_total, 
number_of_people_infected_with_hiv - LAG (number_of_people_infected_with_hiv) 
OVER (ORDER BY hiv_info.Year ASC) 
AS HIV_decrease_cases,
LEAD (number_of_people_infected_with_hiv, 1) OVER (ORDER BY hiv_info.Year ASC) 
AS next_year_HIV_total 
FROM hiv_info INNER JOIN country_info ON Entity = Country AND hiv_info.year = country_info.year
WHERE Country = 'Bermuda';"
df_5.3<-dbGetQuery(mydb2, sql_cd2.5.3)
# check result:
kable(df_5.3)
```

In this query we used the ```LAG``` and ```over``` clause to find out that Bermuda
is keeping a decreasing trend each year on total HIV infected patient. 

#### Query 4

Query number 4, in this query we want to know if the HIV death rate is the same year
with the highest unemployment rate and we use the death rate of HIV per 1000000 people
on different years in Canada.
```{r}
sql_cd2.5.4 <-"SELECT hiv_info.Year, Country, Country_Population, 
CAST(deaths_total * 1000000 AS FLOAT)/CAST(Country_Population AS FLOAT) 
AS 'deaths_rate',GDP_per_capital,unemployment_rate, School_enrollment_rate 
FROM hiv_info INNER JOIN country_info ON Entity = Country 
AND hiv_info.year = country_info.year
WHERE Country = 'Canada' 
ORDER BY (CASE WHEN School_enrollment_rate IS NULL 
THEN unemployment_rate 
ELSE unemployment_rate 
END);"
df_5.4<-dbGetQuery(mydb2, sql_cd2.5.4)
# check result:
kable(df_5.4)
```

We want to order the output table by ```School_enrollment_rate``` but if it is null
then we will alternatively order by ```unemployment_rate```(and since we knew there
are null value in school enrollment rate so we know it will order by unemployment rate)
.As we can see that there is some null value in the ```School_enrollment_rate``` so it
is ordered by ```unemployment_rate``` instead with the help of ```CASE WHEN``` clause.
The highest death rate was on 2007 but we actually have the lowest unemployment rate this
year, so the HIV death rate might have no huge impact on the unemployment rate since they
are not in the same year.  

#### Query 5.

In this query we want to find out the rank of ```new_cases_of_hiv_infection``` each 
year and partition it by country:
```{r}
sql_cd2.5.5 <-"select Country, hiv_info.Year,Country_Population,new_cases_of_hiv_infection , 
dense_rank() over (partition by Country order by new_cases_of_hiv_infection DESC) 
AS 'rank' 
FROM hiv_info INNER JOIN country_info ON Entity = Country 
AND hiv_info.year = country_info.year;"
df_5.5<-dbGetQuery(mydb2, sql_cd2.5.5)
# check result:
kable(df_5.5)
```

The ```partition by``` helps us to get a rank within same country but different 
years and ```dense_rank()``` clause is a window function, which ranking in descending 
ordered partition of ```new_cases_of_hiv_infection```. From the result we can see that 
the Bermuda has the most ```new_cases_of_hiv_infection``` on 2010, 2011, 2012, 2013; 
Canada reached peak at 2017; North America reaches peak at 2017; and US has the 
most new cases on 2017. However, I think the new cases of HIV is increasing accordingly 
with the country population and we can actually use ggplot2 to visualization our result.

```{r}
ggplot(df_5.5, aes(x = Year, y = new_cases_of_hiv_infection, 
                  color="new hiv cases per year" )) +
    geom_line()+
  geom_line(aes(y=Country_Population, color="country population per year")) + 
  scale_x_continuous(breaks=seq(2007, 2017, 1))+
  facet_wrap(. ~ Country)+
  theme_bw()
```

So, we can see that basically for every countries but Bermuda have a increasing 
trending of total population and their HIV new cases are relatively have no change 
in trend at all. Bermuda have a overlapped result but given the fact that its total 
population is not huge and each year of new HIV cases has no more than 10 we still 
can say that all of these countries have a ideal HIV new cases each year. 
The HIV disease in these countries is under control. 

```{r}
dbDisconnect(mydb2)
```
